Spike spectra for recurrences

### 2022-07-26 Manuscript/inter_spike_manuscript_mdpi.tex
Potential restructuring of Fig.1

Some of these recurrence quantification measures, the entropy of diagonal lines and the entropy of 
recurrence times, can be related to the Kolmogorov-Sinai entropy \cite{march2005,baptista2010}. However, these quantifiers have a free parameter, the minimal considered line length, and 
are usually biased, due to the finite size of the RP and thickened diagonal lines, which needs to be corrected \cite{Kraemer2019}. Moreover, the mentioned statistics cannot account for 
changing regular (non-chaotic) dynamics, such as period-doubling bifurcations.
A rather simple idea is to look at the $\tau$-recurrence rate of the RP ($\tau$-RR, Eq.~\ref{eq_tau_rr}) \cite{marwan2002pla,Zbilut2008}.
This is the density of recurrence points along the diagonals of the recurrence matrix, as a function of the distance $\tau$ (sampling units) to the main diagonal:
\begin{equation}\label{eq_tau_rr}
\tau\text{-RR}(\varepsilon) = RR(\tau, \varepsilon) = \frac{1}{N-\tau} \sum_{i=1}^{N-\tau	} R_{i,i+\tau}.
\end{equation}
$\tau$-RR serves as an estimator for the probability that the system recurs after time $\tau \Delta t$, with $\Delta t$ being the sampling time of the trajectory 
$\vec{x}_i = \vec{x}(\Delta t \cdot i),\, i=1,\ldots,N$. 
\citet{Zbilut2008} pointed out that $\tau$-RR could be used as a plugin value for the auto-correlation function $C(\tau)$ and, hence, via the Wiener-Khinchim theorem a 
``generalized'' powerspectrum can be obtained. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and is also preserved in its thresholded version $\tau$-RR. There are clear advantages for a recurrence-derived 
powerspectrum, i.e., Fourier transforming (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D), instead 
of $C(\tau)$ (Fig.~\ref{fig_tau_rr_spectrum_example}C): There are no assumptions for stationarity or sampling, when constructing a RP.
Furthermore, the correlation structures of higher dimensional spaces can be resolved in the recurrence-derived Fourier-spectrum.\\

\begin{figure}
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_spectrum_example}
 \caption{Schematic illustration of a $\tau$-recurrence rate based spectrum. \textbf{A} $x$-component time series of the Lorenz63-System (Eq.~\eqref{eq_model_Lorenz63}) and 
 \textbf{B} its corresponding Fourier power spectrum. 
 \textbf{C} Reconstructed state space portrait from the time series shown in \textbf{A} using PECUZAL time-delay embedding \cite{Kraemer2021}. 
 \textbf{D} Subset of the recurrence plot and the corresponding $\tau$-recurrence rate obtained from the state space trajectory in \textbf{C}. 
 \textbf{E} Fourier Power spectrum obtained from the $\tau$-recurrence rate (subset shown in panel \textbf{D}) \cite{Zbilut2008}. \textbf{D} and \textbf{E} show the results of a part of the time series, 
 which is highlighted in pink in \textbf{A}.\todo[inline]{MA: I think Fig.~\ref{fig_tau_rr_spectrum_example} should be restructured. We could talk about it.\\ HK: I will do that.}
 }\label{fig_tau_rr_spectrum_example}
\end{figure}

However, there is a drawback to this approach. Whenever $\tau$-RR is a spike-train-like signal, which it is in most cases (see Fig.~\ref{fig_tau_rr_spectrum_example}) especially for 
map-data (low-resolution data), a FT of such a signal leads to a spike-train-like image in the frequency domain (e.g. \cite{Schild1982,Cordoba1989}, see 
Fig.~\ref{fig_tau_rr_spectrum_example}D). Thus, it is not intuitive how to extract meaningful information about dominant frequencies of the systems' state space trajectory.

@@

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_spectra}
 \caption{\todo[inline]{could you please become sure that A and B are not the same plots. They are super similar}.Averaged Fourier power spectra of recorded power grid frequency time series of \textbf{A} Central Europe (CE) and \textbf{B} Great Britain (GB) (see Appendix~\ref{sec_power_grid_appendix}). 
 The corresponding averaged inter spike spectra of the according $\tau$-RR's, Eq.~\eqref{eq_tau_rr}, are shown in \textbf{C} and \textbf{D}, respectively. Vertical red dashed lines correspond to 
 $15$, $30$ and $60$ \si{min}. For technical details on the calculation of the spectra shown, the reader is referred to the Appendix~\ref{sec_power_grid_appendix}.}  
\label{fig_power_grid_spectra}
\end{figure}

>>>>

Some of these recurrence quantification measures, the entropy of diagonal lines and the entropy of 
recurrence times, can be related to the Kolmogorov-Sinai entropy \cite{march2005,baptista2010}. However, these quantifiers have a free parameter, the minimal considered line length, and 
are usually biased, due to the finite size of the RP and thickened diagonal lines, which needs to be corrected \cite{Kraemer2019}. Moreover, the mentioned statistics cannot account for 
changing regular (non-chaotic) dynamics, such as period-doubling bifurcations.
A rather simple idea is to look at the $\tau$-recurrence rate of the RP ($\tau$-RR, Eq.~\ref{eq_tau_rr}) \cite{marwan2002pla,Zbilut2008}.
This is the density of recurrence points along the diagonals of the recurrence matrix, as a function of the distance $\tau$ (sampling units) to the main diagonal:
\begin{equation}\label{eq_tau_rr}
\tau\text{-RR}(\varepsilon) = RR(\tau, \varepsilon) = \frac{1}{N-\tau} \sum_{i=1}^{N-\tau	} R_{i,i+\tau}.
\end{equation}
$\tau$-RR serves as an estimator for the probability that the system recurs after time $\tau \Delta t$, with $\Delta t$ being the sampling time of the trajectory 
$\vec{x}_i = \vec{x}(\Delta t \cdot i),\, i=1,\ldots,N$. 
\citet{Zbilut2008} pointed out that $\tau$-RR could be used as a plugin value for the auto-correlation function $C(\tau)$ and, hence, via the Wiener-Khinchim theorem a 
``generalized'' powerspectrum can be obtained. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and is also preserved in its thresholded version $\tau$-RR. There are clear advantages for a recurrence-derived 
powerspectrum, i.e., Fourier transforming (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D,E), instead 
of $C(\tau)$: There are no assumptions for stationarity or sampling, when constructing a RP.
Furthermore, the correlation structures of higher dimensional spaces can be resolved in the recurrence-derived Fourier-spectrum.\\

\begin{figure}
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_spectrum_example}
 \caption{Schematic illustration of a $\tau$-recurrence rate based spectrum. \textbf{A} $x$-component time series of the Lorenz63-System (Eq.~\eqref{eq_model_Lorenz63}) and 
 \textbf{B} its corresponding Fourier power spectrum. 
 \textbf{C} Reconstructed state space portrait from the time series shown in \textbf{A} using PECUZAL time-delay embedding \cite{Kraemer2021}. 
 \textbf{D} Subset of the recurrence plot and the corresponding $\tau$-recurrence rate obtained from the state space trajectory in \textbf{C}. 
 \textbf{E} Fourier Power spectrum obtained from the $\tau$-recurrence rate (subset shown in panel \textbf{D}) \cite{Zbilut2008}. \textbf{D} and \textbf{E} show the results of a part of the time series, 
 which is highlighted in pink in \textbf{A}.
 }\label{fig_tau_rr_spectrum_example}
\end{figure}

However, there is a drawback to this approach. Whenever $\tau$-RR is a spike-train-like signal, which it is in most cases (see Fig.~\ref{fig_tau_rr_spectrum_example}) especially for 
map-data (low-resolution data), a FT of such a signal leads to a spike-train-like image in the frequency domain (e.g. \cite{Schild1982,Cordoba1989}, see 
Fig.~\ref{fig_tau_rr_spectrum_example}E). Thus, it is not intuitive how to extract meaningful information about dominant frequencies of the systems' state space trajectory.

@@

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_spectra}
 \caption{Averaged Fourier power spectra of recorded power grid frequency time series of \textbf{A} Central Europe (CE) and \textbf{B} Great Britain (GB) (see Appendix~\ref{sec_power_grid_appendix}). 
 The corresponding averaged inter spike spectra of the according $\tau$-RR's, Eq.~\eqref{eq_tau_rr}, are shown in \textbf{C} and \textbf{D}, respectively. Vertical red dashed lines correspond to 
 $15$, $30$ and $60$ \si{min}. For technical details on the calculation of the spectra shown, the reader is referred to the Appendix~\ref{sec_power_grid_appendix}.}  
\label{fig_power_grid_spectra}
\end{figure}
